material and methods

### 2017-08-17 man/ms_1_material_methods.Rmd
update structure ms
add content about site and species

---
title: "material and methods"
author: "AJ Perez-Luque (@ajpelu); G. Gea-Izquierdo"
date: "2017"
output:  
  md_document:
    toc: true
    variant: markdown_github
bibliography: ../references.bib
csl: ../ecology.csl
---

@@

## Study sites

Aqui meter descripción de SN y de la distribución del Qp en Sierra Nevada.

@@

>>>>

---
title: "material and methods"
author: "AJ Perez-Luque (@ajpelu); G. Gea-Izquierdo"
date: "2017"
output:  
  md_document:
    toc: true
    variant: markdown_github
bibliography: ../references.bib
csl: ../geb.csl
---

@@

## Species and study site

The Pyrenean oak (*Quercus pyrenaica* Willd.) forests extend through south-western France and the Iberian Peninsula [@Franco1990] reaching its southern limit in north of Morocco. In the Iberian Peninsula these forests live under meso-supramediterranean and mesotemperate areas and subhumid, humid and hyperhumid ombroclimate [@RivasMartinez2002] living on siliceous soils, or soils poor in basic ions [@Vilches2014]. *Q. pyrenaica* requires between 650 and 1200 mm of annual precipitation and a summer minimal precipitation between 100 and 200 mm (Martínez-Parras and Molero-Mesa 1982, García and Jiménez 2009), with summer rainfall being a key factor in the distribution of the species (Gavilán et al. 2007, Río et al. 2007).

This species reaches its southernmost European limit at Sierra Nevada, a high-mountain range located in southern Spain (37°N, 3°W) with altitudes of between 860 m and 3482 m a.s.l. The climate is Mediterranean, characterized by cold winters and hot summers, with pronounced summer drought (July-August). There are eight oak patches (2400 Has) identified (:red_circle: FIGURE) in this mountain range, ranging between 1100 and 2000 *m a.s.l.* and generally associated to major river valleys. Sierra Nevada is considered a glacial refugia for deciduous *Quercus* species during glaciation (Brewer et al. 2002, Olalde et al. 2002, Rodríguez-Sánchez et al. 2010) and these populations are considered as a rear edge of the habitat distribution, which is important in determining habitat responses to expected climate change (Hampe and Petit 2005).

The populations of Pyrenean oak forests at Sierra Nevada are considered relict forests (Melendo and Valle 2000, Vivero et al. 2000), undergoing intensive anthropic use in the last few decades (Camacho-Olmedo et al. 2002, Valbuena-Carabaña et al. 2010). In fact, the status of conservation of this species for southern Spain is "Vulnerable" (Vivero et al. 2000). The relict presence of this species in Sierra Nevada is related both to its genetic resilience as well as to its high intraspecific genetic diversity (Valbuena-Carabaña and Gil 2013). However, they are also expected to suffer the impact of climate change, due to their climate requirements (wet summers). Thus, simulations of the climate change effects on this habitat forecast a reduction in suitable habitats for Sierra Nevada (Benito et al. 2011).

:red_circle: La figura 1 puede tener un mapa de localización de SN, otro de las poblaciones de roble (clasificadas por colores: cluster; y señalando las dos poblaciones muestradas en dendro). Ver MIGRAME dataset

@@

Resilience
